# Modules

**Learning objectives:**

Learn about *modules* with focus on `nn_linear()`, `nn_squential()`, and `nn_module()`

## Built-in `nn_module()s` {.unnumbered}

**What are modules?**

:   -   an object that encapsulates state
    -   can be of any complexity (e.g. layer, or models consisting of layers)

Examples of `{torch}` modules:

-   linear: `nn_linear()`
-   convolutional: `nn_linear()`, `nn_conf1d()`, `nn_conv_3d()`
-   recurrent: `nn_lstm()`, `nn_gru()`
-   embedding: `nn_embedding()`
-   multi-head attention: `nn_multihead_attention()`
-   See [torch documentation](https://torch.mlverse.org/docs/reference/#neural-network-modules) for others

Consider the [linear layer](https://torch.mlverse.org/docs/reference/nn_linear):

```{r}
library(torch)

l <- nn_linear(in_features = 5, out_features = 16) #bias = TRUE is default
l
```

Comment about size: We expect `l` to be $5 \times 16$ (i.e for matrix multiplication: $X_{50\times5}* \beta_{5 \times 16}$. We see below that it is $16 \times 5$, which is due to the underlying C++ implementation of `libtorch`. For performance reasons, the transpose is stored.

```{r}
l$weight$size()
```

Apply the module:

```{r}
#Generate data: generated from the normal distribution
x <- torch_randn(50, 5) 

# Feed x into layer:
output <- l(x)

output$size()
```

When we use built-in modules, `requires_grad = TRUE` is [*not*]{.underline} required in creation of the tensor (unlike previous chapters). It's taken care of for us.

## Sequential Models {.unnumbered}

[`nn_squential()`](https://torch.mlverse.org/docs/reference/nn_sequential) can be used for models consisting solely of linear layers (i.e. a Multi-Layer Perceptron (MLP)). Below we build an MLP using this method:

```{r}
mlp <- nn_sequential( # all arguments should be modules
  nn_linear(10, 32),
  nn_relu(),
  nn_linear(32,64),
  nn_relu(),
  nn_linear(64,1)
)
```

Apply this model to random data:

```{r}
mlp(torch_randn(5, 10))
```

## Non-sequential Models {.unnumbered}

[`nn_module()`](https://torch.mlverse.org/docs/reference/nn_module) is "factory function" for building models of arbitrary complexity. More flexible than the sequential model. Use to define:

-   weight initialization

-   model structure (forward pass), including identification of model parameters using `nn_parameter()` .

Example:

```{r}
my_linear <- nn_module(
  initialize = function(in_features, out_features){
    self$w <- nn_parameter(torch_randn(in_features, out_features)) # random normal
    self$b <- nn_parameter(torch_zeros(out_features))              # zeros
  },
  forward = function(input){
    input$mm(self$w) + self$b
  }
)
```

Next instantiate the model with input and output dimensions:

```{r}
l <- my_linear(7, 1)
l
```

##  {.unnumbered}

## Meeting Videos {.unnumbered}

### Cohort 1 {.unnumbered}

`r knitr::include_url("https://www.youtube.com/embed/URL")`

<details>

<summary>Meeting chat log</summary>

```         
LOG
```

</details>
